---
title: "BDB Visuals"
author: "Pranav Ram"
date: "12/28/2021"
output: html_document
---

```{r setup, include=FALSE}
#load libraries
#install.packages("nflfastR")
library(dplyr)
library(nflfastR)
library(gsisdecoder)
library(ggimage)
library(teamcolors)
library(gt)
library(reactable)
library(reactablefmtr)
```


```{r, include = FALSE}
#Load in Trevor's final ensemble file
full.preds <- read.csv("~/Documents/Documents/R Projects/full preds.txt")
#Drop X column
full.preds = full.preds %>% select(-"X")

#Create new returnTeam column to implement logos
full.preds['returnerTeam'] <- NA
full.preds$returnerTeam <- ifelse(full.preds$possessionTeam == full.preds$homeTeamAbbr, full.preds$visitorTeamAbbr, full.preds$homeTeamAbbr)

#Change all OAK to LV
full.preds$possessionTeam = ifelse(full.preds$possessionTeam == "OAK", "LV", full.preds$possessionTeam)
full.preds$returnerTeam = ifelse(full.preds$returnerTeam == "OAK", "LV", full.preds$returnerTeam)
full.preds$visitorTeamAbbr = ifelse(full.preds$visitorTeamAbbr == "OAK", "LV", full.preds$visitorTeamAbbr)
full.preds$homeTeamAbbr = ifelse(full.preds$homeTeamAbbr == "OAK", "LV", full.preds$homeTeamAbbr)

#Define rounding function
#specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))


```

# First Visual: Table for player ranking with team & Table for player ranking without team
```{r Player Ranking, echo = FALSE}

#Find best & worst returners with Team
bestworst1 <- full.preds %>% 
  group_by(returnerId, returnerTeam, punt_returner_player_name) %>%
  summarise(n = n(), actual_yards = mean(kickReturnYardage), expected_yards = mean(ENS_AVGpreds), yards_over_expected = mean(yard_dif)) %>%
  filter(n >= 10) %>%
  arrange(desc(yards_over_expected))

bestworst1$actual_yards <- round(bestworst1$actual_yards, 2)
bestworst1$expected_yards <- round(bestworst1$expected_yards, 2)
bestworst1$yards_over_expected <- round(bestworst1$yards_over_expected, 2)
#trying to get it uniform, can't get it yet, tried this but it messed up the sorting
#bestworst1$yards_over_expected <- format(round(bestworst1$yards_over_expected, 2), nsmall = 2)
# join team logos to dataset
bestworst1 <- bestworst1 %>%
  left_join(teams_colors_logos, by = c('returnerTeam' = 'team_abbr'))
bestworst1 <- bestworst1 %>%
  select(returnerId, returnerTeam, team_logo_espn, punt_returner_player_name, n, actual_yards, expected_yards, yards_over_expected)


#Plot
reactable(
  bestworst1,
  outlined = TRUE, borderless = TRUE, striped = TRUE, showSortable = TRUE,
  defaultPageSize = 16,
  paginationType = "simple",
  language = reactableLang(
    searchPlaceholder = "Search...",
    noData = "No entries found",
    pageInfo = "{rowStart} to {rowEnd} of {rows} entries",
    pagePrevious = "\u276e",
    pageNext = "\u276f",
    pagePreviousLabel = "Previous page",
    pageNextLabel = "Next page"
  ),
  theme = reactableTheme(
    headerStyle = list(
      "&:hover[aria-sort]" = list(background = "hsl(0, 0%, 96%)"),
      "&[aria-sort='ascending'], &[aria-sort='descending']" = list(background = "hsl(0, 0%, 96%)"),
      borderColor = "#555")),
  columns = list(
    n = colDef(name = "Returns"),
    returnerTeam = colDef(show = FALSE),
    punt_returner_player_name = colDef(show = FALSE),
    actual_yards = colDef(name = "Avg Actual Yards"),
    expected_yards = colDef(name = "Avg Expected Yards"),
    yards_over_expected = colDef(name = "Avg Yards Over Expected",
                                 cell = color_tiles(
                                  bestworst1,
                                  colors = c("#d7191c", "#ffffbf", "#1a9641"),
                                  number_fmt = scales::number_format(accuracy = 0.01))),
    team_logo_espn = colDef(name = "Player", cell = embed_img(bestworst1, label = "punt_returner_player_name")),
    returnerId = colDef(show = FALSE)
  )
)

#Problems so far
#Might not be able to summarize by team as some returners left their teams midseason:
#see Adam Humphries?

################################################################################################

#Find best & worst returners (without Team)
bestworst2 <- full.preds %>% 
  group_by(returnerId, punt_returner_player_name) %>%
  summarise(n = n(), actual_yards = mean(kickReturnYardage), expected_yards = mean(ENS_AVGpreds), yards_over_expected = mean(yard_dif)) %>%
  filter(n >= 10) %>%
  arrange(desc(yards_over_expected))

bestworst2$actual_yards <- format(round(bestworst2$actual_yards, 2), nsmall = 2)
options(digits=2)
bestworst2$actual_yards <- as.numeric(bestworst2$actual_yards)
  #round(bestworst2$actual_yards, 2) 

bestworst2$expected_yards <- format(round(bestworst2$expected_yards, 2), nsmall = 2)
options(digits=2)
bestworst2$expected_yards <- as.numeric(bestworst2$expected_yards)
  #round(bestworst2$expected_yards, 2)

bestworst2$yards_over_expected <- format(round(bestworst2$yards_over_expected, 2), nsmall = 2)
options(digits=2)
bestworst2$yards_over_expected <- as.numeric(bestworst2$yards_over_expected)
  #round(bestworst2$yards_over_expected, 2)

#trying to get it uniform, can't get it yet, tried this but it messed up the sorting
#bestworst1$yards_over_expected <- format(round(bestworst1$yards_over_expected, 2), nsmall = 2)
# join team logos to dataset
bestworst2 <- bestworst2 %>%
  select(returnerId, punt_returner_player_name, n, actual_yards, expected_yards, yards_over_expected)

reactable(
  bestworst2,
  outlined = TRUE, borderless = TRUE, striped = TRUE, showSortable = TRUE,
  defaultPageSize = 16,
  paginationType = "simple",
  language = reactableLang(
    searchPlaceholder = "Search...",
    noData = "No entries found",
    pageInfo = "{rowStart} to {rowEnd} of {rows} entries",
    pagePrevious = "\u276e",
    pageNext = "\u276f",
    pagePreviousLabel = "Previous page",
    pageNextLabel = "Next page"
  ),
  theme = reactableTheme(
    headerStyle = list(
      "&:hover[aria-sort]" = list(background = "hsl(0, 0%, 96%)"),
      "&[aria-sort='ascending'], &[aria-sort='descending']" = list(background = "hsl(0, 0%, 96%)"),
      borderColor = "#555")),  columns = list(
    n = colDef(name = "Returns"),
    punt_returner_player_name = colDef(name = "Player"),
    actual_yards = colDef(name = "Avg Actual Yards"),
    expected_yards = colDef(name = "Avg Expected Yards"),
    yards_over_expected = colDef(name = "Avg Yards Over Expected",
                                 cell = color_tiles(
                                  bestworst2,
                                  colors = c("#d7191c", "#ffffbf", "#1a9641"),
                                  number_fmt = scales::number_format(accuracy = 0.01))),
    returnerId = colDef(show = FALSE)

  )
)


```

# Second Visual: Table for team punt coverage ranking
```{r Cover Team Ranking, echo=FALSE}
#Find best & worst teams with getting in position
coverteam1 <- full.preds %>% 
  group_by(possessionTeam) %>% 
  summarise(n = n(), yards_given_up = mean(kickReturnYardage), yards_expected = mean(ENS_AVGpreds), yards_expected_diff = mean(yard_dif)) %>%
  arrange(yards_expected)
coverteam1$yards_expected <- round(coverteam1$yards_expected, 2)
coverteam1$yards_given_up <- round(coverteam1$yards_given_up, 2)
coverteam1$yards_expected_diff <- round(coverteam1$yards_expected_diff, 2)
# join team logos to dataset
coverteam1 <- coverteam1 %>%
  left_join(teams_colors_logos, by = c('possessionTeam' = 'team_abbr'))
coverteam1 <- coverteam1 %>%
  select(possessionTeam, team_logo_espn, n, yards_given_up, yards_expected, yards_expected_diff)


#Plot
reactable(
  coverteam1,
  outlined = TRUE, borderless = TRUE, striped = TRUE, showSortable = TRUE,
  defaultPageSize = 16,
  paginationType = "simple",
  language = reactableLang(
    searchPlaceholder = "Search...",
    noData = "No entries found",
    pageInfo = "{rowStart} to {rowEnd} of {rows} entries",
    pagePrevious = "\u276e",
    pageNext = "\u276f",
    pagePreviousLabel = "Previous page",
    pageNextLabel = "Next page"
  ),
  theme = reactableTheme(
    headerStyle = list(
      "&:hover[aria-sort]" = list(background = "hsl(0, 0%, 96%)"),
      "&[aria-sort='ascending'], &[aria-sort='descending']" = list(background = "hsl(0, 0%, 96%)"),
      borderColor = "#555")),  columns = list(
    possessionTeam = colDef(show = FALSE),
    n = colDef(maxWidth = 100, name = "Returns"),
    yards_given_up = colDef(name = "Yards Allowed Per Return"),
    yards_expected_diff = colDef(name = "Average Difference",
                                 cell = color_tiles(
                                  coverteam1,
                                  colors = c("#1a9641", "#ffffbf", "#d7191c"),
                                  number_fmt = scales::number_format(accuracy = 0.01))),
    yards_expected = colDef(name = "Expected Yards Allowed Per Return"),
    team_logo_espn = colDef(maxWidth = 100, name = "Team", cell = embed_img(coverteam1, label = "possessionTeam"))
      )
)

```


# Third Visual: Table for team punt blocking ranking
```{r Return Team Ranking, echo = FALSE}
#Find best & worst teams with getting in position
returnteam1 <- full.preds %>% 
  group_by(returnerTeam) %>% 
  summarise(n = n(), actual_return_yards = mean(kickReturnYardage), yards_expected = mean(ENS_AVGpreds), yards_expected_diff = mean(yard_dif)) %>%
  arrange(desc(yards_expected))
returnteam1$yards_expected <- round(returnteam1$yards_expected, 2)
returnteam1$actual_return_yards <- round(returnteam1$actual_return_yards, 2)
returnteam1$yards_expected_diff <- round(returnteam1$yards_expected_diff, 2)
# join team logos to dataset
returnteam1 <- returnteam1 %>%
  left_join(teams_colors_logos, by = c('returnerTeam' = 'team_abbr'))
returnteam1 <- returnteam1 %>%
  select(returnerTeam, team_logo_espn, n, actual_return_yards, yards_expected, yards_expected_diff)

#Plot
reactable(
  returnteam1,
  outlined = TRUE, borderless = TRUE, striped = TRUE, showSortable = TRUE,
  defaultPageSize = 16,
  paginationType = "simple",
  language = reactableLang(
    searchPlaceholder = "Search...",
    noData = "No entries found",
    pageInfo = "{rowStart} to {rowEnd} of {rows} entries",
    pagePrevious = "\u276e",
    pageNext = "\u276f",
    pagePreviousLabel = "Previous page",
    pageNextLabel = "Next page"
  ),
  theme = reactableTheme(
    headerStyle = list(
      "&:hover[aria-sort]" = list(background = "hsl(0, 0%, 96%)"),
      "&[aria-sort='ascending'], &[aria-sort='descending']" = list(background = "hsl(0, 0%, 96%)"),
      borderColor = "#555")),  columns = list(
    returnerTeam = colDef(show = FALSE),
    n = colDef(maxWidth = 100, name = "Returns"),
    actual_return_yards = colDef(name = "Yards Per Return"),
    yards_expected_diff = colDef(name = "Average Difference",
                                 cell = color_tiles(
                                  returnteam1,
                                  colors = c("#d7191c", "#ffffbf" ,"#1a9641"),
                                  number_fmt = scales::number_format(accuracy = 0.01))),
    yards_expected = colDef(name = "Expected Yards Per Return"),
    team_logo_espn = colDef(maxWidth = 100, name = "Team", cell = embed_img(returnteam1, label = "returnerTeam"))
  )
)

```

# Testing to see if we can Nest Teams with an aggregate to see who the best returners are for each team
```{r Test, echo = FALSE}
#Nesting by Team
#Find best & worst returners with Team
bestworst3 <- full.preds %>% 
  group_by(returnerId, returnerTeam, punt_returner_player_name) %>%
  summarise(n = n(), actual_yards = mean(kickReturnYardage), expected_yards = mean(ENS_AVGpreds), yards_over_expected = mean(yard_dif)) %>%
  filter(n >= 10) %>%
  arrange(desc(yards_over_expected))

bestworst3$actual_yards <- round(bestworst3$actual_yards, 2)
bestworst3$expected_yards <- round(bestworst3$expected_yards, 2)
bestworst3$yards_over_expected <- round(bestworst3$yards_over_expected, 2)
#trying to get it uniform, can't get it yet, tried this but it messed up the sorting
#bestworst1$yards_over_expected <- format(round(bestworst1$yards_over_expected, 2), nsmall = 2)
# join team logos to dataset
bestworst3 <- bestworst3 %>%
  left_join(teams_colors_logos, by = c('returnerTeam' = 'team_abbr'))
bestworst3 <- bestworst3 %>%
  select(returnerId, returnerTeam, team_logo_espn, punt_returner_player_name, n, actual_yards, expected_yards, yards_over_expected)


#Plot
reactable(
  bestworst3,
  outlined = TRUE, borderless = TRUE, striped = TRUE, showSortable = TRUE,
  defaultPageSize = 16,
  language = reactableLang(
    searchPlaceholder = "Search...",
    noData = "No entries found",
    pageInfo = "{rowStart} to {rowEnd} of {rows} entries",
    pagePrevious = "\u276e",
    pageNext = "\u276f",
    pagePreviousLabel = "Previous page",
    pageNextLabel = "Next page"
  ),
  groupBy = "returnerTeam",
  columns = list(
    n = colDef(name = "Returns"),
    returnerTeam = colDef(show = FALSE),
    punt_returner_player_name = colDef(show = FALSE),
    actual_yards = colDef(name = "Avg Actual Yards"),
    expected_yards = colDef(name = "Avg Expected Yards"),
    team_logo_espn = colDef(name = "Team", cell = embed_img(bestworst1, label = "punt_returner_player_name")),
    returnerId = colDef(show = FALSE),
    yards_over_expected = colDef(name = "Avg Yards Over Expected",
                                 cell = color_tiles(
                                  bestworst1,
                                  colors = c("#d7191c", "#ffffbf", "#1a9641"),
                                  number_fmt = scales::number_format(accuracy = 0.01)),
                                 aggregate = "sum",
      # Render aggregated cell as currency
      aggregated = JS("function(cellInfo) {
        return cellInfo.value.toFixed(2)
      }")))
)

```
# Will play around with this visual some more and see if I can get it to where it looks like "Visual 3" with a dropdown menu for each team to look an individual returners.