---
title: "Data Preprocessing"
author: "Trevor Hughes"
date: "12/26/2021"
output: html_document
---
```{r}
library(dplyr)
library(stringr)
library(sjmisc)
library(stringr)

```


############################### Data Prep ###############################
#read in data
```{r}
#Returned datasets were created in Data Subsetting.ipynb
returned2018 = read.csv("C:\\Users\\thughes\\Documents\\returned2018.csv")
returned2019 = read.csv("C:\\Users\\thughes\\Documents\\returned2019.csv")
returned2020 = read.csv("C:\\Users\\thughes\\Documents\\returned2020.csv")
```
#Filter data to only the the time stamp where the returner first received the punt
```{r}
received2018 = returned2018 %>% filter(event == "punt_received")
received2019 = returned2019 %>% filter(event == "punt_received")
received2020 = returned2020 %>% filter(event == "punt_received")

received = rbind(received2018, received2019, received2020)
```

#Calculate each players distance from the returner
```{r}
#Indicates which observations have multiple returners on one play
received$multiple_returners = ifelse(str_detect((received$returnerId), ";"), 1, 0)
sum(received$multiple_returners)

received = received %>% filter(multiple_returners == 0)

#Indicates if a player is a returner or not
received$returner = ifelse(received$returnerId == received$nflId, 1, 0)

#Creates separate dataset with the returners location and joins it with each row in the original data set where the game and play id match
receivedReturners = received %>% filter(returner == 1) %>% select(gameId, playId,x, y) %>% rename(returner_x = x, returner_y = y)

received = left_join(received, receivedReturners, by = c("gameId", "playId"))

#Calculates the distance of each player from the returner
received = received %>% mutate(distance = sqrt((returner_x - x)^2 +  (returner_y - y)^2))
```


#Other calculations to standardize
```{r}
#Determine if each player is on the kicking or receiving team
received$playerTeam = ifelse(received$team == "away", received$visitorTeamAbbr, received$homeTeamAbbr)

received$kickingTeam = ifelse(received$playerTeam == received$possessionTeam, 1, 0)

#Indicates if the receiving team is home or away
received$receiveTeamHome = ifelse(received$possessionTeam != received$homeTeamAbbr, 1, 0)

#Score differential in respect to receiving team
received$receiveTeamScoreDif = ifelse(received$possessionTeam != received$homeTeamAbbr, received$preSnapHomeScore - received$preSnapVisitorScore, received$preSnapVisitorScore - received$preSnapHomeScore)

#Calculate overall yardline number
received$yardlineTotal = ifelse(received$yardlineSide != received$possessionTeam, 100 - received$yardlineNumber, received$yardlineNumber)


punters = received %>% filter(nflId == kickerId) %>% select(gameId, playId,x, y) %>% rename(punter_x = x, punter_y = y)

received = left_join(received, punters, by = c("gameId", "playId"))


#Orient x and y axis to always start where 0 is on the receiving end
received$absX = ifelse(received$punter_x > received$returner_x, received$x, 120 - received$x)
received$absY = ifelse(received$punter_x > received$returner_x, received$y, 53.3 - received$y)

#Remove punts that were deflected
received = received %>% filter(!str_detect(playDescription, "deflected"))

#Zone splits
xSplit = 120/5
ySplit = 53.3/3
#Builds a zone grid for player location
received$zone = ifelse(received$absX <= xSplit, ifelse(received$absY <= ySplit, 1, ifelse(received$absY <= 2*ySplit, 2, 3)),  ifelse(received$absX <= 2*xSplit, ifelse(received$absY <= ySplit, 4, ifelse(received$absY <= 2*ySplit, 5, 6)),  ifelse(received$absX <= 3*xSplit, ifelse(received$absY <= ySplit, 7, ifelse(received$absY <= 2*ySplit, 8, 9)), ifelse(received$absX <= 4*xSplit, ifelse(received$absY <= ySplit, 10, ifelse(received$absY <= 2*ySplit, 11, 12)), ifelse(received$absY <= ySplit, 13, ifelse(received$absY <= 2*ySplit, 14, 15))))))

#Creates a variable that indicates which side of the field a defender is on from the defenders perspective
received$side = ifelse(received$zone %in% c(1,4,7,10,13), "L", ifelse(received$zone %in% c(2,5,8,11,14), "C", "R"))
```

Transforming data frame to be modelling friendly 
```{r}
#Create empty data frame to use for modeling
model_set_def = data.frame(gameId = integer(), playId = integer(),
                       def1_dist = integer(), def1_s = integer(),
                       def1_a = integer(), def2_dist = integer(), 
                       def2_s = integer(), def2_a = integer(), 
                       def3_dist = integer(), def3_s = integer(),
                       def3_a = integer(),  def4_dist = integer(),
                       def4_s = integer(), def4_a = integer(), 
                       def5_dist = integer(), def5_s = integer(),
                       def5_a = integer(), def6_dist = integer(),
                       def6_s = integer(), def6_a = integer(), 
                       def7_dist = integer(), def7_s = integer(),
                       def7_a = integer(),  def8_dist = integer(), 
                       def8_s = integer(), def8_a = integer(), 
                       def9_dist = integer(), def9_s = integer(),
                       def9_a = integer(), def10_dist = integer(), 
                       def10_s = integer(), def10_a = integer(), 
                       def11_dist = integer(), def11_s = integer(),
                       def11_a = integer(),def1_zone = integer(), def2_zone = integer(),
                       def3_zone = integer(), def4_zone = integer(), 
                       def5_zone = integer(), def6_zone = integer(), 
                       def7_zone = integer(), def8_zone = integer(),
                       def9_zone = integer(), def10_zone = integer(),
                       def11_zone = integer(), def1_side = character(), 
                       def2_side = character(), def3_side = character(), 
                       def4_side = character(), def5_side = character(), 
                       def6_side = character(), def7_side = character(), 
                       def8_side = character(), def9_side = character(),
                       def10_side = character(), def11_side = character()
                       
                       )

# Loops through each unique game and creates a new dataset with each game
for (games in unique(received$gameId)) {
  x= received %>% filter(gameId == games & kickingTeam == 1)
  
  #Loops through each unique play in the game and creates a new dataset for each play
  for (plays in unique(x$playId)) {
    y = x %>% filter(playId == plays)
    
    #Loops through 11 for each player on the kicking team
    for (i in 1:11) {
      #Finds the shortest distance from a defender to the returner saves the value
      assign(paste0("min", i, sep = ""), min(y$distance, na.rm = TRUE))
      
      #Filters to the player with that shortest distance and creates values for their speed, acceleration, and position
      z = y[which.min(y$distance),]
      assign(paste0("s", i, sep = ""), z$s)
      assign(paste0("a", i, sep = ""), z$a)
      assign(paste0("zone", i, sep = ""), z$zone)
      assign(paste0("side", i, sep = ""), z$side)
  
      #Removes that player from the temporary set and loops again
      y = y[-which.min(y$distance),]
    }
    
    #Prints out all of the values into one data frame
    model_set_def[nrow(model_set_def)+1,] <- c(games, plays, min1, s1, a1, 
                                       min2, s2, a2,
                                       min3, s3, a3,
                                       min4, s4, a4,
                                       min5, s5, a5,
                                       min6, s6, a6,
                                       min7, s7, a7,
                                       min8, s8, a8,
                                       min9, s9, a9, 
                                       min10, s10, a10,
                                       min11, s11, a11,
                                       zone1, zone2, zone3, zone4, zone5, zone6,
                                       zone7, zone8, zone9, zone10, zone11, 
                                       side1, side2, side3, side4, side5, side6, 
                                       side7, side8, side9, side10, side11)
  
  
  }
}

#Convert all distance, acceleration, and speed variables to numeric
names = c(3:35)

model_set_def[,names] = lapply(model_set_def[,names],function(x) as.numeric(as.character(x)))

```


#Creating model set for returner
```{r}
#Create empty data frame to use for modeling
model_set_ret = data.frame(gameId = integer(), playId = integer(), return_x = integer(),
                       return_s = integer(), return_a = integer(),
                       return_zone = integer())

# Loops through each unique game and creates a new dataset with each game
for (games in unique(received$gameId)) {
  x= received %>% filter(gameId == games & returner == 1)
  
  #Loops through each unique play in the game and creates a new dataset for each play
  for (plays in unique(x$playId)) {
    y = x %>% filter(playId == plays)
    
    ret_x = y$absX
    ret_s = y$s
    ret_a = y$a
    ret_zone = y$zone
    
    #Prints out all of the values into one data frame
    model_set_ret[nrow(model_set_ret)+1,] <- c(games, plays, ret_x, ret_s, ret_a, ret_zone)
  }

}

#Convert all distance, acceleration, and speed variables to numeric
names = c(3:5)

model_set_ret[,names] = lapply(model_set_ret[,names],function(x) as.numeric(as.character(x)))

table(model_set_ret$return_zone)

```

#Creating a model set with player position for offense, Transforming data frame to be modelling friendly 
```{r}
#Create empty data frame to use for modeling
model_set_off = data.frame(gameId = integer(), playId = integer(),
                       off1_dist = integer(), off1_s = integer(),
                       off1_a = integer(), off2_dist = integer(), off2_s = integer(),
                       off2_a = integer(), off3_dist = integer(), off3_s = integer(),
                       off3_a = integer(), off4_dist = integer(), off4_s = integer(),
                       off4_a = integer(), off5_dist = integer(), off5_s = integer(),
                       off5_a = integer(), off6_dist = integer(), off6_s = integer(),
                       off6_a = integer(), off7_dist = integer(), off7_s = integer(),
                       off7_a = integer(), off8_dist = integer(), off8_s = integer(),
                       off8_a = integer(), off9_dist = integer(), off9_s = integer(),
                       off9_a = integer(), off10_dist = integer(), off10_s = integer(),
                       off10_a = integer(), off11_dist = integer(), off11_s = integer(),
                       off11_a = integer()
                       )


# Loops through each unique game and creates a new dataset with each game
for (games in unique(received$gameId)) {
  x= received %>% filter(gameId == games & kickingTeam == 0)
  
  #Loops through each unique play in the game and creates a new dataset for each play
  for (plays in unique(x$playId)) {
    y = x %>% filter(playId == plays)
    
    #Loops through 11 for each player on the receiving team
    for (i in 1:11) {
      #Finds the shortest distance from a blocker to the returner saves the value
      assign(paste0("min", i, sep = ""), min(y$distance, na.rm = TRUE))
      
      #Filters to the player with that shortest distance and creates values for their speed, acceleration, and position
      z = y[which.min(y$distance),]
      assign(paste0("s", i, sep = ""), z$s)
      assign(paste0("a", i, sep = ""), z$a)
      assign(paste0("pos", i, sep = ""), z$position)
  
      #Removes that player from the temporary set and loops again
      y = y[-which.min(y$distance),]
    }
    
    #Prints out all of the values into one data frame
    model_set_off[nrow(model_set_off)+1,] <- c(games, plays, as.integer(min1), s1, a1,
                                       min2, s2, a2, 
                                       min3, s3, a3,
                                       min4, s4, a4,
                                       min5, s5, a5, 
                                       min6, s6, a6, 
                                       min7, s7, a7, 
                                       min8, s8, a8,
                                       min9, s9, a9, 
                                       min10, s10, a10, 
                                      min11, s11, a11)

  
  
  }
}

#Convert all distance, acceleration, and speed variables to numeric
names = c(3:35)

model_set_off[,names] = lapply(model_set_off[,names],function(x) as.numeric(as.character(x)))
```

#Add variables from dataset with epa
```{r}
epa2018 = read.csv("C:\\Users\\thughes\\Documents\\epa2018.csv")
epa2019 = read.csv("C:\\Users\\thughes\\Documents\\epa2019.csv")
epa2020 = read.csv("C:\\Users\\thughes\\Documents\\epa2020.csv")
epa = rbind(epa2018, epa2019, epa2020)
epa$gameId = epa$old_game_id
epa$playId = epa$play_id

#Impute wind and temp values for indoor stadiums
epa$wind = ifelse(is.na(epa$wind), 0, epa$wind)
epa$temp = ifelse(is.na(epa$temp), 70, epa$temp)

epa_model_vars = epa%>% select("gameId", "playId", "series", "surface",  "temp", "wind")


epa_full_vars = epa %>% select("gameId", "playId", "weather", "ep", "epa", "punt_returner_player_name", "punter_player_name", "roof", "home_coach", "away_coach")

```



############################### Merge the data sets ####################################
```{r}


#Selects the variables we want to include in model
received_model_vars = received %>% select("gameId", "playId",  "playDirection", "returnDirectionIntended", "returnDirectionActual", "quarter", "yardlineTotal", "receiveTeamScoreDif","kickLength", "kickReturnYardage", "receiveTeamHome")
#, "playResult"

#Sets game and play id to character variables
received_model_vars$gameId = as.character(received_model_vars$gameId)
received_model_vars$playId =as.character(received_model_vars$playId)

#Joins the data sets and doesn't use any duplicated rows for one play
all_players =  merge( model_set_ret, model_set_def, by = c("gameId", "playId"))
final_set =  merge( all_players, received_model_vars[!duplicated(received_model_vars), ], by = c("gameId", "playId"))
final_set = merge(final_set, model_set_off, by = c("gameId", "playId"))
final_set =  merge( final_set, epa_model_vars[!duplicated(epa_model_vars), ], by = c("gameId", "playId"))

#Drop off1 data since we already have returner info
final_set = final_set %>% select(-c(off1_s, off1_a, off1_dist))

#Write out to csvs
write.csv(final_set, "C:\\Users\\thughes\\Documents\\final set.csv")

```

#Create data set with all variables
```{r}


received_full_vars = received %>% select("gameId", "playId", "missedTackler", "tackler", "assistTackler", "vises", "season", "week", "kickContactType", "gameDate", "gameTimeEastern", "homeTeamAbbr", "visitorTeamAbbr", "playDescription", "yardsToGo", "possessionTeam", "kickerId", "returnerId", "gameClock", "penaltyCodes", "penaltyJerseyNumbers", "penaltyYards", "playResult")

received_full_vars$gameId = as.character(received_full_vars$gameId)
received_full_vars$playId =as.character(received_full_vars$playId)

#Joins the data sets and doesn't use any duplicated rows for one play

full_set =  merge( final_set, received_full_vars[!duplicated(received_full_vars), ], by = c("gameId", "playId"))
full_set =  merge( full_set, epa_full_vars[!duplicated(epa_full_vars), ], by = c("gameId", "playId"))


write.csv(full_set, "C:\\Users\\thughes\\Documents\\full set.csv")
```