---
title: "neural_network_final"
author: "Liam Dao"
date: "12/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Setup
```{r import packages, warning = FALSE, message=FALSE}
library(tidyverse)
library(caret)
library(leaps)
library(glmnet)
library(ggplot2)
library(earth)
library(mgcv)
library(ROCR)
library(Cairo)
library(pROC)
library(ROCR)
library(randomForest)
library(xgboost)
library(Ckmeans.1d.dp)
library(pdp)
library(Matrix)
library(gganimate)
library(cowplot)
library(ggridges)
library(repr)
library(gifski)
library(plotly)
library(rpart)
library(rpart.plot)
library(nnet)
library(NeuralNetTools)
library(Metrics)

#turning off warnings
options(warn=-1)

# Set directory
setwd("C:\\Users\\liamd\\Documents\\GitHub\\NFL_Big_Data_Bowl_2022\\data_folder")

# Read in data
train <- read.csv("model train set.csv")
val <- read.csv("model val set.csv")
test <- read.csv("model test set.csv")

train3 <- train[,c(4:40,63:64,66:68,70,71 )]
val3 <- val[,c(4:40,63:64,66:68,70,71 )]
test3 <- test[,c(4:40,63:64,66:68,70,71 )]
```

## Variable Analysis and Data Posturing (Train)
```{r variable analysis}
# Remove def zone variables & actual return direction
train <- train %>% 
  select(-41:-51, -65)

# Any variables with less than 10 values are labeled at categorical (13 because I wanted zone variables to be a factor)
for (i in 1:length(train)) {
  if (length(unique(train[,i])) < 16) {
    train[,i] = as.factor(train[,i])
  }
}

# Filter out NA observations
train = train %>% filter(is.na(train$kickReturnYardage) == 0)

#Getting variable importance from decision tree
fit = rpart(train$kickReturnYardage ~ . , data = train[,c(4:59)])
rpart.plot(fit)
varimp.data = data.frame(fit$variable.importance)
varimp.data$names = as.character(rownames(varimp.data))
varimp.data

# Check columns data types
sapply(train, class) # Dataset contains numeric, integer, and factor variables

# Scale standardize 
for (i in 4:length(train)) {
  if (class(train[,i]) == c('numeric', 'integer')) {
    train[,i] = scale(train[,i])
  }
}

# Using varimp.data output. Cutoff of 700 for fit.variable.importance
var.sel <- varimp.data$names

train2 <- train %>% 
  select(var.sel, kickReturnYardage, X, gameId, playId)
```

## Data Posturing (Validation)
```{r data posturing (validation)}
# Remove def zone variables & actual return direction
val <- val %>% 
  select(-41:-51, -65)

# Any variables with less than 10 values are labeled at categorical (16 because I wanted zone variables to be a factor)
for (i in 1:length(val)) {
  if (length(unique(val[,i])) < 16) {
    val[,i] = as.factor(val[,i])
  }
}

# Filter out NA observations
val <-  val %>% 
  filter(is.na(val$kickReturnYardage) == 0)

# Scale standardize 
for (i in 4:length(val)) {
  if (class(val[,i]) == c('numeric', 'integer')) {
    val[,i] = scale(val[,i])
  }
}

val$kickReturnYardage <- as.numeric(val$kickReturnYardage)

# Subset variables
val2 <- val %>% 
  select(var.sel, X, kickReturnYardage, gameId, playId)
```

## Data Posturing (test)
```{r data posturing (test)}
# Remove def zone variables & actual return direction
test <- test %>% 
  select(-41:-51, -65)

# Any variables with less than 10 values are labeled at categorical (16 because I wanted zone variables to be a factor)
for (i in 1:length(test)) {
  if (length(unique(test[,i])) < 16) {
    test[,i] = as.factor(test[,i])
  }
}

# Filter out NA observations
test <-  test %>% 
  filter(is.na(test$kickReturnYardage) == 0)

# Scale standardize 
for (i in 4:length(test)) {
  if (class(test[,i]) == c('numeric', 'integer')) {
    test[,i] = scale(test[,i])
  }
}

test$kickReturnYardage <- as.numeric(test$kickReturnYardage)

# Subset variables
test2 <- test %>% 
  select(var.sel, X, kickReturnYardage, gameId, playId)
```

## Model Creation
```{r model creation}
set.seed(123)
nn.yard.model <- nnet(kickReturnYardage ~ . -gameId -playId, data = train2, size = 6, decay = .945, linout = TRUE )
```

## Fit on Validation
```{r fit on validation}
nn.yard.val.predict <- predict(nn.yard.model, newdata = val2)

# Calculate RMSE and MAE
val.rmse <- sqrt(mean((nn.yard.val.predict - val2$kickReturnYardage)^2)) # RMSE
val.mae <- mae(val2$kickReturnYardage, nn.yard.val.predict) # MAE
```
> Performance Metrics:
RMSE: 8.33
MAE: 5.28

## Fit on Test
```{r fit on test}
nn.yard.test.predict <- predict(nn.yard.model, newdata = test2)

# Calculate RMSE and MAE
test.rmse <- sqrt(mean((nn.yard.test.predict - test2$kickReturnYardage)^2)) # RMSE
test.mae <- mae(test2$kickReturnYardage, nn.yard.test.predict) # MAE
```
> Performance Metrics:
RMSE: 11.45
MAE: 6.83


############################# TRY WITH ALL DATA (not including zones) ##################################
## Model Creation
```{r model creation}
set.seed(123)
nn.yard.model <- nnet(kickReturnYardage ~ . -gameId -playId, data = train, size = 6, decay = 1, linout = TRUE )
```

## Fit on Validation
```{r fit on validation}
nn.yard.val.predict <- predict(nn.yard.model, newdata = val)

# Calculate RMSE and MAE
val.rmse <- sqrt(mean((nn.yard.val.predict - val$kickReturnYardage)^2)) # RMSE
val.mae <- mae(val2$kickReturnYardage, nn.yard.val.predict) # MAE
```
> Performance Metrics:
RMSE: 8.42
MAE: 5.73

## Fit on Test
```{r fit on test}
nn.yard.test.predict <- predict(nn.yard.model, newdata = test)

# Calculate RMSE and MAE
test.rmse <- sqrt(mean((nn.yard.test.predict - test2$kickReturnYardage)^2)) # RMSE
test.mae <- mae(test$kickReturnYardage, nn.yard.test.predict) # MAE
```
> Performance Metrics:
RMSE: 11.49
MAE: 6.99


############################# TRY WITH 44 VARIABLES (from trevor) ##################################
## Variable Analysis and Data Posturing (Train)
```{r variable analysis}
# Any variables with less than 10 values are labeled at categorical (13 because I wanted zone variables to be a factor)
for (i in 1:length(train3)) {
  if (length(unique(train3[,i])) < 16) {
    train3[,i] = as.factor(train3[,i])
  }
}

# Filter out NA observations
train3 = train3 %>% filter(is.na(train3$kickReturnYardage) == 0)

# Scale standardize 
for (i in 4:length(train3)) {
  if (class(train3[,i]) == c('numeric', 'integer')) {
    train3[,i] = scale(train3[,i])
  }
}
```

## Data Posturing (Validation)
```{r data posturing (validation)}
# Any variables with less than 10 values are labeled at categorical (13 because I wanted zone variables to be a factor)
for (i in 1:length(val3)) {
  if (length(unique(val3[,i])) < 16) {
    val3[,i] = as.factor(val3[,i])
  }
}

# Filter out NA observations
val3 = val3 %>% filter(is.na(val3$kickReturnYardage) == 0)

# Scale standardize 
for (i in 4:length(val3)) {
  if (class(val3[,i]) == c('numeric', 'integer')) {
    val3[,i] = scale(val3[,i])
  }
}
```

## Data Posturing (test)
```{r data posturing (test)}
# Any variables with less than 10 values are labeled at categorical (13 because I wanted zone variables to be a factor)
for (i in 1:length(test3)) {
  if (length(unique(test3[,i])) < 16) {
    test3[,i] = as.factor(test3[,i])
  }
}

# Filter out NA observations
test3 = test3 %>% filter(is.na(test3$kickReturnYardage) == 0)

# Scale standardize 
for (i in 4:length(test3)) {
  if (class(test3[,i]) == c('numeric', 'integer')) {
    test3[,i] = scale(test3[,i])
  }
}
```

## Model Creation
```{r model creation}
set.seed(123)
nn.yard.model3 <- nnet(kickReturnYardage ~ . , data = train3, size = 3, decay = .35, linout = TRUE )
```

## Fit on Validation
```{r fit on validation}
nn.yard.val3.predict <- predict(nn.yard.model3, newdata = val3)

# Calculate RMSE and MAE
val3.rmse <- sqrt(mean((nn.yard.val3.predict - val3$kickReturnYardage)^2)) # RMSE
val3.mae <- mae(val3$kickReturnYardage, nn.yard.val3.predict) # MAE
```
> Performance Metrics:
RMSE: 8.03
MAE: 5.39

## Fit on Test
```{r fit on test}
nn.yard.test3.predict <- predict(nn.yard.model3, newdata = test3)

# Calculate RMSE and MAE
test3.rmse <- sqrt(mean((nn.yard.test3.predict - test3$kickReturnYardage)^2)) # RMSE
test3.mae <- mae(test3$kickReturnYardage, nn.yard.test3.predict) # MAE
```
> Performance Metrics:
RMSE: 11.37
MAE: 6.81