---
title: "Offense_Players_BDB"
author: "Pranav Ram"
date: "12/21/2021"
output: html_document
---

############################### Data Cleaning #################################
#Setup
```{r setup, include=FALSE}
#Libraries
library(dplyr)
library(stringr)
library(sjmisc)
library(stringr)

#Local File
#Same as the received 18,19,20 files binded together in Trevor's Code
received = read.csv('BDB_Final')
#I consolidated all of these to a csv on my laptop to make it easier

```

#Trevor's Code to Posture
```{r }
#Indicates which observations have multiple returners on one play
received$multiple_returners = ifelse(str_detect((received$returnerId), ";"), 1, 0)
sum(received$multiple_returners)

received = received %>% filter(multiple_returners == 0)

#Indicates if a player is a returner or not
received$returner = ifelse(received$returnerId == received$nflId, 1, 0)

#Creates separate dataset with the returners location and joins it with each row in the original data set where the game and play id match
receivedReturners = received %>% filter(returner == 1) %>% select(gameId, playId,x, y) %>% rename(returner_x = x, returner_y = y)

received = left_join(received, receivedReturners, by = c("gameId", "playId"))

#Calculates the distance of each player from the returner
received = received %>% mutate(distance = sqrt((returner_x - x)^2 +  (returner_y - y)^2))

#Determine if each player is on the kicking or receiving team
received$playerTeam = ifelse(received$team == "away", received$visitorTeamAbbr, received$homeTeamAbbr)

received$kickingTeam = ifelse(received$playerTeam == received$possessionTeam, 1, 0)

#Indicates if the receiving team is home or away
received$receiveTeamHome = ifelse(received$possessionTeam != received$homeTeamAbbr, 1, 0)

#Score differential in respect to receiving team
received$receiveTeamScoreDif = ifelse(received$possessionTeam != received$homeTeamAbbr, received$preSnapHomeScore - received$preSnapVisitorScore, received$preSnapVisitorScore - received$preSnapHomeScore)

#Calculate overall yardline number
received$yardlineTotal = ifelse(received$yardlineSide != received$possessionTeam, 100 - received$yardlineNumber, received$yardlineNumber)


punters = received %>% filter(nflId == kickerId) %>% select(gameId, playId,x, y) %>% rename(punter_x = x, punter_y = y)

received = left_join(received, punters, by = c("gameId", "playId"))

#Orient x and y axis to always start where 0 is on the receiving end
received$absX = ifelse(received$punter_x > received$returner_x, received$x, 120 - received$x)
received$absY = ifelse(received$punter_x > received$returner_x, received$y, 53.3 - received$y)

#Remove punts that were deflected
received = received %>% filter(!str_detect(playDescription, "deflected"))

#Zone splits
xSplit = 120/5
ySplit = 53.3/3
#Builds a zone grid for player location
received$zone = ifelse(received$absX <= xSplit, ifelse(received$absY <= ySplit, 1, ifelse(received$absY <= 2*ySplit, 2, 3)),  ifelse(received$absX <= 2*xSplit, ifelse(received$absY <= ySplit, 4, ifelse(received$absY <= 2*ySplit, 5, 6)),  ifelse(received$absX <= 3*xSplit, ifelse(received$absY <= ySplit, 7, ifelse(received$absY <= 2*ySplit, 8, 9)), ifelse(received$absX <= 4*xSplit, ifelse(received$absY <= ySplit, 10, ifelse(received$absY <= 2*ySplit, 11, 12)), ifelse(received$absY <= ySplit, 13, ifelse(received$absY <= 2*ySplit, 14, 15))))))

#Creates a variable that indicates which side of the field a defender is on from the defenders perspective
received$side = ifelse(received$zone %in% c(1,4,7,10,13), "L", ifelse(received$zone %in% c(2,5,8,11,14), "C", "R"))

```

Transforming data frame to be modelling friendly (Including player position)
```{r}
#Create empty data frame to use for modeling
model_set = data.frame(gameId = integer(), playId = integer(),
                       def1_dist = integer(), def1_s = integer(),
                       def1_a = integer(), def1_pos = character(),
                       def2_dist = integer(), def2_s = integer(),
                       def2_a = integer(), def2_pos = character(),
                       def3_dist = integer(), def3_s = integer(),
                       def3_a = integer(), def3_pos = character(),
                       def4_dist = integer(), def4_s = integer(),
                       def4_a = integer(), def4_pos = character(),
                       def5_dist = integer(), def5_s = integer(),
                       def5_a = integer(), def5_pos = character(),
                       def6_dist = integer(), def6_s = integer(),
                       def6_a = integer(), def6_pos = character(),
                       def7_dist = integer(), def7_s = integer(),
                       def7_a = integer(), def7_pos = character(),
                       def8_dist = integer(), def8_s = integer(),
                       def8_a = integer(), def8_pos = character(),
                       def9_dist = integer(), def9_s = integer(),
                       def9_a = integer(), def9_pos = character(),
                       def10_dist = integer(), def10_s = integer(),
                       def10_a = integer(), def10_pos = character(),
                       def11_dist = integer(), def11_s = integer(),
                       def11_a = integer(), def11_pos = character()
                       )

# Loops through each unique game and creates a new dataset with each game
for (games in unique(received$gameId)) {
  x= received %>% filter(gameId == games & kickingTeam == 1)
  
  #Loops through each unique play in the game and creates a new dataset for each play
  for (plays in unique(x$playId)) {
    y = x %>% filter(playId == plays)
    
    #Loops through 11 for each player on the kicking team
    for (i in 1:11) {
      #Finds the shortest distance from a defender to the returner saves the value
      assign(paste0("min", i, sep = ""), min(y$distance, na.rm = TRUE))
      
      #Filters to the player with that shortest distance and creates values for their speed, acceleration, and position
      z = y[which.min(y$distance),]
      assign(paste0("s", i, sep = ""), z$s)
      assign(paste0("a", i, sep = ""), z$a)
      assign(paste0("pos", i, sep = ""), z$position)
  
      #Removes that player from the temporary set and loops again
      y = y[-which.min(y$distance),]
    }
    
    #Prints out all of the values into one data frame
    model_set[nrow(model_set)+1,] <- c(games, plays, as.integer(min1), s1, a1, pos1,
                                       min2, s2, a2, pos2,
                                       min3, s3, a3, pos3,
                                       min4, s4, a4, pos4,
                                       min5, s5, a5, pos5,
                                       min6, s6, a6, pos6,
                                       min7, s7, a7, pos7,
                                       min8, s8, a8, pos8,
                                       min9, s9, a9, pos9,
                                       min10, s10, a10, pos10,
                                       min11, s11, a11, pos11)
  
  
  }
}

#Convert all distance, acceleration, and speed variables to numeric
names = c(3:5, 7:9, 11:13, 15:17, 19:21, 23:25, 27:29, 31:33, 35:37, 39:41, 43:45)

model_set[,names] = lapply(model_set[,names],function(x) as.numeric(as.character(x)))
```

#Creating a model set with player position for offense, Transforming data frame to be modelling friendly (Including player position)
```{r}
#Create empty data frame to use for modeling
model_set2 = data.frame(gameId = integer(), playId = integer(),
                       off1_dist = integer(), off1_s = integer(),
                       off1_a = integer(), off1_pos = character(),
                       off2_dist = integer(), off2_s = integer(),
                       off2_a = integer(), off2_pos = character(),
                       off3_dist = integer(), off3_s = integer(),
                       off3_a = integer(), off3_pos = character(),
                       off4_dist = integer(), off4_s = integer(),
                       off4_a = integer(), off4_pos = character(),
                       off5_dist = integer(), off5_s = integer(),
                       off5_a = integer(), off5_pos = character(),
                       off6_dist = integer(), off6_s = integer(),
                       off6_a = integer(), off6_pos = character(),
                       off7_dist = integer(), off7_s = integer(),
                       off7_a = integer(), off7_pos = character(),
                       off8_dist = integer(), off8_s = integer(),
                       off8_a = integer(), off8_pos = character(),
                       off9_dist = integer(), off9_s = integer(),
                       off9_a = integer(), off9_pos = character(),
                       off10_dist = integer(), off10_s = integer(),
                       off10_a = integer(), off10_pos = character(),
                       off11_dist = integer(), off11_s = integer(),
                       off11_a = integer(), off11_pos = character()
                       )



# Loops through each unique game and creates a new dataset with each game
for (games in unique(received$gameId)) {
  x= received %>% filter(gameId == games & kickingTeam == 0)
  
  #Loops through each unique play in the game and creates a new dataset for each play
  for (plays in unique(x$playId)) {
    y = x %>% filter(playId == plays)
    
    #Loops through 11 for each player on the receiving team
    for (i in 1:11) {
      #Finds the shortest distance from a blocker to the returner saves the value
      assign(paste0("min", i, sep = ""), min(y$distance, na.rm = TRUE))
      
      #Filters to the player with that shortest distance and creates values for their speed, acceleration, and position
      z = y[which.min(y$distance),]
      assign(paste0("s", i, sep = ""), z$s)
      assign(paste0("a", i, sep = ""), z$a)
      assign(paste0("pos", i, sep = ""), z$position)
  
      #Removes that player from the temporary set and loops again
      y = y[-which.min(y$distance),]
    }
    
    #Prints out all of the values into one data frame
    model_set2[nrow(model_set2)+1,] <- c(games, plays, as.integer(min1), s1, a1, pos1,
                                       min2, s2, a2, pos2,
                                       min3, s3, a3, pos3,
                                       min4, s4, a4, pos4,
                                       min5, s5, a5, pos5,
                                       min6, s6, a6, pos6,
                                       min7, s7, a7, pos7,
                                       min8, s8, a8, pos8,
                                       min9, s9, a9, pos9,
                                       min10, s10, a10, pos10,
                                      min11, s11, a11, pos11)

  
  
  }
}

#Convert all distance, acceleration, and speed variables to numeric
names = c(3:5, 7:9, 11:13, 15:17, 19:21, 23:25, 27:29, 31:33, 35:37, 39:41, 43:45)

model_set2[,names] = lapply(model_set2[,names],function(x) as.numeric(as.character(x)))
```


#Creating model set without player position for defense
```{r }
#Create empty data frame to use for modeling
model_set_def = data.frame(gameId = integer(), playId = integer(),
                       def1_dist = integer(), def1_s = integer(),
                       def1_a = integer(), def2_dist = integer(), 
                       def2_s = integer(), def2_a = integer(), 
                       def3_dist = integer(), def3_s = integer(),
                       def3_a = integer(),  def4_dist = integer(),
                       def4_s = integer(), def4_a = integer(), 
                       def5_dist = integer(), def5_s = integer(),
                       def5_a = integer(), def6_dist = integer(),
                       def6_s = integer(), def6_a = integer(), 
                       def7_dist = integer(), def7_s = integer(),
                       def7_a = integer(),  def8_dist = integer(), 
                       def8_s = integer(), def8_a = integer(), 
                       def9_dist = integer(), def9_s = integer(),
                       def9_a = integer(), def10_dist = integer(), 
                       def10_s = integer(), def10_a = integer(), 
                       def11_dist = integer(), def11_s = integer(),
                       def11_a = integer(),def1_zone = integer(), def2_zone = integer(),
                       def3_zone = integer(), def4_zone = integer(), 
                       def5_zone = integer(), def6_zone = integer(), 
                       def7_zone = integer(), def8_zone = integer(),
                       def9_zone = integer(), def10_zone = integer(),
                       def11_zone = integer(), def1_side = character(), 
                       def2_side = character(), def3_side = character(), 
                       def4_side = character(), def5_side = character(), 
                       def6_side = character(), def7_side = character(), 
                       def8_side = character(), def9_side = character(),
                       def10_side = character(), def11_side = character()
                       
                       )

# Loops through each unique game and creates a new dataset with each game
for (games in unique(received$gameId)) {
  x= received %>% filter(gameId == games & kickingTeam == 1)
  
  #Loops through each unique play in the game and creates a new dataset for each play
  for (plays in unique(x$playId)) {
    y = x %>% filter(playId == plays)
    
    #Loops through 11 for each player on the kicking team
    for (i in 1:11) {
      #Finds the shortest distance from a defender to the returner saves the value
      assign(paste0("min", i, sep = ""), min(y$distance, na.rm = TRUE))
      
      #Filters to the player with that shortest distance and creates values for their speed, acceleration, and position
      z = y[which.min(y$distance),]
      assign(paste0("s", i, sep = ""), z$s)
      assign(paste0("a", i, sep = ""), z$a)
      assign(paste0("zone", i, sep = ""), z$zone)
      assign(paste0("side", i, sep = ""), z$side)
  
      #Removes that player from the temporary set and loops again
      y = y[-which.min(y$distance),]
    }
    
    #Prints out all of the values into one data frame
    model_set_def[nrow(model_set_def)+1,] <- c(games, plays, min1, s1, a1, 
                                       min2, s2, a2,
                                       min3, s3, a3,
                                       min4, s4, a4,
                                       min5, s5, a5,
                                       min6, s6, a6,
                                       min7, s7, a7,
                                       min8, s8, a8,
                                       min9, s9, a9, 
                                       min10, s10, a10,
                                       min11, s11, a11,
                                       zone1, zone2, zone3, zone4, zone5, zone6,
                                       zone7, zone8, zone9, zone10, zone11, 
                                       side1, side2, side3, side4, side5, side6, 
                                       side7, side8, side9, side10, side11)
  
  
  }
}

#Convert all distance, acceleration, and speed variables to numeric
names = c(3:35)

model_set_def[,names] = lapply(model_set_def[,names],function(x) as.numeric(as.character(x)))
```


#Creating model set without player position for offense
```{r}
#Create empty data frame to use for modeling
model_set_off = data.frame(gameId = integer(), playId = integer(), return_x = integer(),
                       return_s = integer(), return_a = integer(),
                       return_zone = integer())

# Loops through each unique game and creates a new dataset with each game
for (games in unique(received$gameId)) {
  x= received %>% filter(gameId == games & returner == 1)
  
  #Loops through each unique play in the game and creates a new dataset for each play
  for (plays in unique(x$playId)) {
    y = x %>% filter(playId == plays)
    
    ret_x = y$absX
    ret_s = y$s
    ret_a = y$a
    ret_zone = y$zone
    
    #Prints out all of the values into one data frame
    model_set_off[nrow(model_set_off)+1,] <- c(games, plays, ret_x, ret_s, ret_a, ret_zone)
  }

}

#Convert all distance, acceleration, and speed variables to numeric
names = c(3:5)

model_set_off[,names] = lapply(model_set_off[,names],function(x) as.numeric(as.character(x)))

check = received %>% filter(gameId == 2018101100)

table(model_set_off$return_zone)

```


#Merge the rest of the important variables
```{r}


#Selects the variables we want to include in model
received_model_vars = received %>% select("gameId", "playId",  "playDirection", "returnDirectionIntended", "returnDirectionActual", "quarter", "yardlineTotal", "receiveTeamScoreDif","kickLength", "kickReturnYardage", "receiveTeamHome")
#, "playResult"

#Sets game and play id to character variables
received_model_vars$gameId = as.character(received_model_vars$gameId)
received_model_vars$playId =as.character(received_model_vars$playId)

model_set_offense = merge( model_set2, model_set_off, by = c("gameId", "playId"))

#Joins the data sets and doesn't use any duplicated rows for one play
all_players =  merge( model_set_offense, model_set_def, by = c("gameId", "playId"))
final_set =  merge( all_players, received_model_vars[!duplicated(received_model_vars), ], by = c("gameId", "playId"))
#final_set =  merge( final_set, epa_model_vars[!duplicated(epa_model_vars), ], by = c("gameId", "playId"))
#couldn't add EPA because didn't have wifi on the plane

final_set = final_set %>% select(-c(off1_s, off1_a, off1_dist, off1_pos, off2_pos, off3_pos, off4_pos, off5_pos, off6_pos, off7_pos, off8_pos , off9_pos, off10_pos, off11_pos))

set.seed(123)
train = final_set %>% sample_frac(.7)
non_train = anti_join(final_set, train, by = c('gameId', 'playId'))
set.seed(123)
val = non_train %>% sample_frac(.2)
test = anti_join(non_train, val, by = c('gameId', 'playId'))

write.csv(final_set, "model set.csv")
write.csv(train, "train set.csv")
write.csv(val, "val set.csv")
write.csv(test, "model test set.csv")

```


############################### Modeling #################################

```{r}
#Change all variables with less than 16 unique values to factor
for (i in 1:length(train)) {
  if (length(unique(train[,i])) < 16) {
    train[,i] = as.factor(train[,i])
  }
}

train = train %>% filter(is.na(train$kickReturnYardage) == 0)

for (i in 1:length(val)) {
  if (length(unique(val[,i])) < 16) {
    val[,i] = as.factor(val[,i])
  }
}
  
val = val %>% filter(is.na(val$kickReturnYardage) == 0)

for (i in 1:length(full_set)) {
  if (length(unique(full_set[,i])) < 16) {
    full_set[,i] = as.factor(full_set[,i])
  }
}
```


```{r}
class(final_set$def1_dist)
lm1 = lm(kickReturnYardage ~ . -gameId -playId, data = final_set)
summary(lm1)
#What we got goin on?
```

offense and defense:
- distance, speed and acceleration for all 21 players (minus the returner)
others:
- playDirection
- returnDirectionIntended
- returnDirectionActual
- quarter
- yardlineTotal
- receiveTeamScoreDif
- kickLength
- receiveTeamHome


```{r}
 #Included only listed variables
  train_x = model.matrix(kickReturnYardage ~ ., 
                         data = train[,c(3:69,92:100 )])[,-1]
  train_y = train$kickReturnYardage
  
  tune_grid = expand.grid(
    nrounds = 4,
    eta = c(.1, .15, .2, .25, .3),
    max_depth = c(1:10),
    gamma = c(0),
    colsample_bytree = 1,
    min_child_weight = 1,
    subsample = c(.25, .5, .75, 1))
```


```{r}
# Build XGBoost model
set.seed(123)
xgb <- xgb.cv(data = train_x, label = train_y, subsample = 0.5, nrounds = 100, nfold = 10, eval.metric = 'mae')
#[4]	test MAE lowest at nrounds = 4

## Caret Tuning
#set seed
set.seed(1)
# Tuning through caret
tune_grid <- expand.grid(
  nrounds = 10,
  eta = c(0.1, 0.2, 0.3),
  max_depth = c(1:5),
  gamma = c(0),
  colsample_bytree = 1,
  min_child_weight = 1,
  subsample = c(0.5, 1)
)

#set seed
set.seed(123)
#build caret tuning
xgb.punt.caret = train(x = train_x, y = train_y, method = "xgbTree", 
                         tuneGrid = tune_grid, 
                         trControl = trainControl(method = 'cv', number = 10),
                         metric = 'MAE')

#plot tuning grid
plot(xgb.punt.caret)
#4 rounds
#0.3 eta
#subsample of 1.00

xgb.punt.jamaica <- xgb.cv(data = train_x, label = train_y, subsample = 1, nrounds = 100, nfold = 10, eta = 0.3, max_depth = 4, eval.metric = 'mae')
#best at 6?

############################################################################################################
xgb.punt.jamaica <- xgboost(data = train_x, label = train_y, subsample = 1, nrounds = 4, eta = 0.3, max_depth = 4, eval.metric = 'mae')
#MAE = 5.114753

xgb.punt.jamaica <- xgboost(data = train_x, label = train_y, subsample = 1, nrounds = 6, eta = 0.3, max_depth = 4, eval.metric = 'mae')
#MAE = 4.838582
############################################################################################################

```

#Try it on validation
```{r}



```

